\Title

### 2022-09-29 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.
Introduction: 1.2

A new generation of algorithms has emerged with the advent of the so-called silicon retina, which mimics the output of the retina by representing information using spikes. %A novel emerging representation is that provided by event-based cameras, in which each pixel independently processes its input and emits an event for positive or negative increments of the log-luminance. 
The shift from the classical, dense representations to this sparse encoding of visual information offers a better analogy to neurobiology, but also offers more energy-efficient computations.  In a recent study, we have introduced a spiking neural network as a classification model applied to event streams based on Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022} which extended the HOTS model~\citet{lagorce_hots_2017}. This HOTS model represents temporal relationships between events by building ``time surfaces'', 2D images computed using the time difference to the last recorded events. By transforming each event in the stream as a vectorial input, our MLR classifier was able to make a decision for every single event, that is in an online fashion. We have demonstrated on several datasets that it provides with efficient computations resulting in ultra-fast categorization. Additionally, we made a formal bridge between this event-based MLR and a SNN, demonstrating the bio-plausibility of this method and its possible integration to neuromorphic hardware. A series of similar algorithms are applied to real-life applications such as optical flow~\citep{benosman_asynchronous_2012}, depth estimation~\citep{dardelet_event-by-event_2021}, or gesture recognition~\citep{maro_event-based_2020}. Recently, such an algorithm has been used in the space industry which is famous for its stringent selection process.

These computational results suggest that the temporal representation underlying spiking representation may be a key ingredient in the efficiency of the algorithms, but one may wonder how this can be supported by neurobiological data. In 1982, Abeles asked if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking motifs~\citep{abeles_role_1982}. While the first possibility favors the rate coding theory, the second highlights the function of temporal precision in the neural code. Since, numerous studies demonstrated the emergence of synchronicity in the activity within a neural population~\citep{riehle_spike_1997, davis_spontaneous_2021}, efficient encoding thanks to the use of spike latencies~\citep{perrinet_coding_2004, gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}. All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of repeated spatio-temporal motifs in biological spike trains. In neuronal models, an efficient use or detection of these spatio-temporal motifs embedded in the spike train comes with the integration of heterogeneous delays~\citep{gutig_tempotron_2006, guise_bayesian_2014, zhang_supervised_2020}. Notably,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, relative spiking delays. This representation has a much greater information capacity in comparison to a firing-rate based neural coding approach through the variety of configurations and the possible coexistence of multiple superposed motifs.

>>>>

A new generation of algorithms has emerged with the advent of so-called silicon retinas, which mimics the output of the retina by representing visual information using spikes. %A novel emerging representation is that provided by event-based cameras, in which each pixel independently processes its input and emits an event for positive or negative increments of the log-luminance. 
The shift from the classical, dense representations to this sparse encoding of visual information offers a better analogy to neurobiology, but also offers more energy-efficient computations. The HOTS model~\citet{lagorce_hots_2017} represents temporal relationships between events by building ``time surfaces'', 2D images computed using the time difference to the last recorded events. In a recent study, we have introduced a spiking neuron network as a classification model applied to event-based streams based on Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022} which extended the HOTS model. By transforming each event in the stream as a vectorial input, our MLR classifier was able to make a decision for every single event, that is, in an online fashion. We have demonstrated on several benchmark datasets that it provides with efficient computations resulting in ultra-fast categorization. Additionally, we made a formal bridge between this event-based MLR and a SNN, demonstrating the bio-plausibility of this method and its possible integration to neuromorphic hardware. A series of similar algorithms are applied to real-life applications such as optical flow~\citep{benosman_asynchronous_2012}, depth estimation~\citep{dardelet_event-by-event_2021}, or gesture recognition~\citep{maro_event-based_2020}. Recently, such an algorithm was used in the CASPR mission on the ISS, illustrating, given the stringent selection processes used by the space sector, that such algorithms have reached maturity for industry.

These computational results suggest that the temporal representation underlying processing in spiking neurons may be a key ingredient in the efficiency of the algorithms, but one may wonder how this can be supported by neurobiological data. In 1982, Abeles asked if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking motifs~\citep{abeles_role_1982}. While the first possibility favors the rate coding theory, the second highlights the function of temporal precision in the neural code. Since, numerous studies demonstrated the emergence of synchronicity in the activity within a neural population~\citep{riehle_spike_1997, davis_spontaneous_2021}, efficient encoding thanks to the use of spike latencies~\citep{perrinet_coding_2004, gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}. All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of spatio-temporal spiking motifs in biological spike trains. In neuronal models, an efficient use or detection of these spatio-temporal motifs embedded in the spike train comes with the integration of heterogeneous delays~\citep{gutig_tempotron_2006, guise_bayesian_2014, zhang_supervised_2020}. Notably,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, relative spiking delays. This representation has a much greater information capacity in comparison to a firing-rate based neural coding approach through the variety of configurations and the possible coexistence of multiple superposed motifs.
